---
title: "ESM 206 Lab 5"
author: "Elliott Matthews"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 objectives

- Learn to parse dates with 'lubridate'
- Get counts of observations with count()
- Uncount() rows
- One and two sample t-test
- Create a heatmap with geom_tile()


```{r, include = FALSE}

# Read in data and attach packages

library(tidyverse)
library(here)
library(janitor)

# REPLACING NA CODE CLEANUP BELOW##########################

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"),
                              na = "-99999") %>% 
  clean_names()
  

```

Use the tidyr::uncount() function to convert our lobster data from frequency format to case format.

```{r}
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)

# uncount seperated counted lobster column into individual values
  
```

#### Exploratory data visualization

We're only going to consider 'site' as our variable of interest

```{r}

ggplot(lobster_tidy, aes(x= site, y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2)

# Histogram
# facet_wrap(~site, scales = "free") will show data on different x and y scales making them look more like each other overall and less easy to objectively compare.

ggplot(lobster_tidy, aes(x = size_mm))+
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

# qq plots wants an argument that says a sample is this for geom qq

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)
```

#### Convert the 'date' column to class 'Date'

We'll use the 'lubridate' package to convert to Date format, and then to help us easily parse month & year

```{r}
lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date))


```

Now lets parse year and month using lubridate::month() and lubridate::year():

```{r}
lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate:: year(date_new))

    # label = TRUE in month() changes the numeric value for month to the TLA
    # it is an ordered factor.  Order is Jan - Dec
  

```

Now lets find counts of observed lobsters based on different hieracrhial groupings:

First: let's count lobsters by year and month:

```{r}
lobster_ym <- lobster_parse_date %>% 
  count(obs_year,obs_month)  # this counts as count observations with (this & this)

lobster_ym

lobster_y <- lobster_parse_date %>% 
  count(obs_year)

lobster_y

lobster_site <- lobster_parse_date %>% 
  count(site)

lobster_site
```

If we want to create a summary table that contains statistics OTHER than counts by group, it's easier to use group_by + n()

```{r}

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    lobster_number = n()
  )

lobster_summary

##### Use group_by if you want to make a summary table of multiple different counts
##### Otherwise 'count' is simpler and quicker for single counts.
```

#### Find confidence intervals

Use t.test() function to find confidence intervals (for one sample) and perform t-tests to compare means of two samples (...this will be convered conceptually in lectures Week 6)

```{r}
ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobsters)

```


#### Two-sample t-test to compare means

We want to ask: Is there a significant difference in lobster lengths at Naples and Mohawk reefs?

We've done our necessary exploratory analyses to determine that a 2-sample t-test for means comparison is appropriate.

```{r}
napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site== "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, napl_sample)

mn_ttest

```

There is a significant difference in lobster lengths between Naples and Mohawk Reef (t(df) = statistic, p < 0.001, alpha =0.05).

```{r}

lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <- t.test(size_mm ~ site, data = lobster_mn)

mn_ttest2

mn_ttest2$p.value
mn_ttest2$statistic

```

Here is the p-value: `r mn_ttest2$p.value`

# WHOA ^^^^^ That's purty neet.

Here is my t-statistic:`r mn_ttest2$statistic`

#### Now a heatmap!

```{r}

lobster_ys <- lobster_parse_date %>% 
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))

```

